To Be Decided

### 2025-07-24 paper/main.tex
Introduction written

        \IEEEPARstart{H}{ere} we go.

        % O que é Microwave Imaging.

        % O papel dos algoritmos de inversão dentro de MWI para resolver o problema de espalhamento eletromagnético e quais desafios (ill-posedness etc). Destacar também a formulação 2D.

@@

@@

@@

@@

@@

		Let $D \in \mathbb{R}^2$ denote the Domain of Interest (DOI) embedded within a homogeneous, isotropic, nonmagnetic ($\mu = \mu_0 = 4\pi \times 10^{-7}$ H/m), and lossless ($\sigma = \sigma_0 = 0$ S/m) background medium with permittivity $\epsilon_b = \epsilon_{rb}\epsilon_0$, where $\epsilon_0 \approx 8.85 \times 10^{-12}$ F/m. We consider a 2D Transverse Magnetic (TMz) polarization where the DOI is illuminated by incident plane waves and assume the time convention $e^{j\omega t}$. All scatterers are located within the DOI, and the scattered field $E^s_z$ at point $\brho \in S$ outside $D$ is evaluated according to the following integral equation \cite{harrington2001time}:
		\begin{equation}
			E^s_z (\brho) = -\frac{k_b^2}{4} \int_D H_0^{(2)}(k_b|\brho-\brhop|)\chi(\brhop)E_z(\brhop) dS^\prime
		\end{equation}

@@

        The experiments were conducted using the open-source library EISPY2D \cite{batista2025eispy2d}, specifically developed to design and evaluate algorithms for the inverse electromagnetic scattering problem. Both indicators will be analyzed, with each approach featuring two case studies (one fundamental and the other exploring a particular aspect) as well as a benchmarking study. The case studies aim to assess performance in specific scenarios, while the benchmarking study facilitates broader generalization of the results.

@@

        For all experiments, the following common parameters were used: background wavelength $\lambda_b = 1$ m, noise level of 5\%, relative permittivity of the background medium $\epsilon_{rb} = 1$, and incident wave amplitude of 1 V/m.

@@

                The DNL distribution is shown in Fig. \ref{fig:position:average:dnl}. The values range from below 1 to above 4, with a median close to 2. The high contrast level and geometric variations contributed to the wide distribution of DNL values. However, as shown in Fig. \ref{fig:position:average:zeta_s}, the algorithm performance within each instance remained within a similar range. CSI clearly exhibited the worst performance, followed by SOM. The best performance in each instance varied between OSM and CA. These results are consistent with the findings from Section \ref{sec:results:position:single}.

@@

@@

@@

>>>>

        % O Imageamento em Microondas é uma importante técnica para a detecção e caracterização de objetos em meios inacessíveis \cite{pastorino2010microwave}. Essa técnica de baixo custo utiliza ondas eletromagnéticas de baixa potência para obter informações sobre a estrutura interna de materiais de maneira não invasiva \cite{godinho2025evaluation}. As aplicações deste método são diversas, incluindo diagnósticos médicos \cite{mojabi2025microwave}, inspeção de estruturas \cite{fedeli2025mild}, through-wall imaging \cite{fedeli2020through}, sensoreamento remoto \cite{salucci2017multifrequency}, entre outros.

        \IEEEPARstart{M}{icrowave} Imaging is an important technique for the detection and characterization of objects in inaccessible media \cite{pastorino2010microwave}. This low-cost technique uses low-power electromagnetic waves to obtain information about the internal structure of materials in a non-invasive manner \cite{godinho2025evaluation}. The applications of this method are diverse, including medical diagnostics \cite{mojabi2025microwave}, structural inspection \cite{fedeli2025mild}, through-wall imaging \cite{fedeli2020through}, remote sensing \cite{salucci2017multifrequency}, among others.

        % Esta técnica é baseada na solução do problema inverso de espalhamento eletromagnético \cite{chen2018computational}. Ou seja, a partir das medições do campo eletromagnético fora do domínio de interesse, busca-se recuperar as propriedades do meio interno. Este problema é conhecido por ser mal posto, o que significa que pequenas variações nas medições podem levar a grandes variações na solução \cite{colton2019inverse}. Além disso, ele também é não-linear e não-convexo devido aos efeitos de espalhamento múltiplo \cite{chen2018computational}.

        This technique is based on solving the inverse electromagnetic scattering problem \cite{chen2018computational}. Specifically, electromagnetic field measurements collected outside the Domain of Interest (DOI) are used to recover the properties of the internal medium. This problem is inherently ill-posed, meaning that small variations in measurements can lead to large variations in the solution \cite{colton2019inverse}. Furthermore, the problem is both nonlinear and non-convex due to multiple scattering effects \cite{chen2018computational}. The problem can be formulated in either two or three dimensions, with two-dimensional formulations being more commonly used for algorithm development due to their computational simplicity and efficiency.

@@

        % Existe uma ampla gama de algoritmos para resolver o problema inverso de espalhamento eletromagnético. Estes algoritmos podem ser classificados em duas categorias principais: métodos qualitativos \cite{pastorino2010ch5}, que reconstroem a imagem sem estimar as propriedades elétricas do meio, e métodos quantitativos \cite{pastorino2010ch7}, os quais estimam essas propriedades. Entre os métodos qualitativos podemos destacar o Linear Sampling Method (LSM) \cite{chen2017lsm}, o Orthogonality Sampling Method (OSM) \cite{harris2020orthogonality}, entre outros. Entre os métodos quantitativos, podemos destacar o Born Iterative Method (BIM) \cite{wang1989iterative}, o Contrast Source Inversion (CSI) \cite{berg1997contrast}, o Distorted Born Inversion Method (DBIM) \cite{chew1990reconstruction}, o Subspace Optimization Method (SOM) \cite{chen2010subspace}, entre outros. Vale à pena destacar também que há algoritmos que contemplam certas especificidades do problema: como o método DORT \cite{chen2018computational} que aborda a reversão temporal dos sinais eletromagnéticos, o MUSIC \cite{zhong2007music}, o qual é adequado para pequenos espalhadores, o Compressive Sensing \cite{massa2015compressive}, o qual pode ser aplicado quando os dados de campo não tem fase, a Aproximação de Born de Primeira Ordem, o Backpropagation e a aproximação de Rytov \cite{chen2018computational}, os quais são úteis para problemas com espalhadores fracos. Outras abordagens relevantes são os Experimentos Virtuais \cite{isernia2024some,bevacqua2021simple}, os quais recombinam os dados do campo espalhado para explorar condições que aliviam a não-linearidade, e a reescrita eficiente das equações \cite{isernia2024some,bevacqua2021effective}, as quais tem o mesmo objetivo final de aliviar a não linearidade. Vale à pena destacar inclusive que, recentemente, algoritmos que fazem o uso de aprendizado de máquina tem sido propostos \cite{chen2020review}. Dentre as abordagens existentes, os métodos orientados a dados (data-driven), como o Direct Inversion Scheme \cite{yao2020enhanced,wei2019solving}, permitem a inferência direta dos parâmetros físicos de um espalhador a partir dos campos medidos, minimizando a necessidade de pré-processamento. Contudo, alguns métodos frequentemente demandam vastos conjuntos de dados para treinamento e apresentam notórias limitações em termos de generalização, confiabilidade e interpretabilidade. Para mitigar essas deficiências, foram propostos os métodos orientados à física (physics-driven), que incorporam leis e informações físicas para guiar o processo de reconstrução \cite{wei2019physics,wang2023push,hu2023more,liu2025exploring}. Outra abordagem é a construção de modelos substitutos para recuperação da posição, geometria e contraste dos espalhadores a partir de um problema de otimização \cite{salucci2022learned,zardi2025physics}. Apesar de superarem algumas das falhas das abordagens puramente baseadas em dados, sua eficácia pode estar condicionada a parâmetros do sistema, como a função de Green, cuja obtenção em meios complexos ou inomogêneos representa um desafio significativo.

        A wide range of algorithms exists for solving the inverse electromagnetic scattering problem. These algorithms can be broadly classified into two main categories: qualitative methods \cite{pastorino2010ch5}, which reconstruct images without estimating the electrical properties of the medium, and quantitative methods \cite{pastorino2010ch7}, which estimate these properties.

        Among qualitative methods, notable examples include the Linear Sampling Method (LSM) \cite{chen2017lsm} and the Orthogonality Sampling Method (OSM) \cite{harris2020orthogonality}. Among quantitative methods, prominent approaches include the Born Iterative Method (BIM) \cite{wang1989iterative}, Contrast Source Inversion (CSI) \cite{berg1997contrast}, the Distorted Born Inversion Method (DBIM) \cite{chew1990reconstruction}, and the Subspace Optimization Method (SOM) \cite{chen2010subspace}.

        Specialized algorithms have been developed to address specific problem characteristics. These include the DORT method \cite{chen2018computational} for electromagnetic signal time reversal, MUSIC \cite{zhong2007music} for small scatterers, and Compressive Sensing \cite{massa2015compressive} for phaseless field data. For weak scatterer problems, effective approaches include First-Order Born Approximation, Backpropagation, and Rytov approximation \cite{chen2018computational}.

        Additional relevant techniques include Virtual Experiments \cite{isernia2024some,bevacqua2021simple}, which recombine scattered field data to explore conditions that alleviate nonlinearity, and efficient equation reformulations \cite{isernia2024some,bevacqua2021effective} that share the same objective of reducing nonlinearity.

        Recently, machine learning-based algorithms have emerged as promising solutions \cite{chen2020review}. Among these approaches, data-driven methods such as the Direct Inversion Scheme \cite{yao2020enhanced,wei2019solving} enable direct inference of scatterer physical parameters from measured fields, minimizing preprocessing requirements. However, these methods often demand extensive training datasets and exhibit significant limitations in generalization, reliability, and interpretability.

        To address these deficiencies, physics-driven methods have been proposed that incorporate physical laws and information to guide the reconstruction process \cite{wei2019physics,wang2023push,hu2023more,liu2025exploring}. Another approach involves constructing surrogate models for recovering scatterer position, geometry, and contrast through optimization problems \cite{salucci2022learned,zardi2025physics}. Despite overcoming some limitations of purely data-driven approaches, their effectiveness may depend on system parameters such as the Green's function, whose determination in complex or inhomogeneous media represents a significant challenge.

@@

        % Em grande parte destes trabalhos, os algoritmos propostos são testados e avaliados em métricas como o mean square error (MSE) da estimativa do contraste pixel a pixel ou algo similar \cite{yin2023subspace,zhang2022023,salucci2022learned,wang2023push,liu2022som,bevacqua2021simple,bevacqua2021effective}. Este tipo de abordagem leva em consideração a precisão na estimativa do contraste na forma direta e, indiretamente, pode dar uma perspectiva de quão boa a reconstrução da posição e geometria dos objetos pode ser, uma vez que erros na geometria e posicionamento podem afetar nessa métrica. No entanto, uma geometria bem precisa com uma má estimativa do contraste pode resultar em um erro considerável neste tipo de métrica, por exemplo. Logo, aspectos comparar o desempenho no aspecto de quão bem a geometria e a posição dos objetos são recuperadas não são contemplados adequadamente por este tipo de métrica. Além disso, a métrica MSE não é adequada para avaliar algoritmos qualitativos, uma vez que estes não estimam o contraste do meio. Vale à pena destacar também a métrica Structural Similarity (SSIM) \cite{wang2004image} to quantitatively show the reconstruction visual quality, which is designed to compute the mismatch between reconstructed contrast results and corresponding GTs both on the pixel-wise level and perceptual level. Ao invés de fazer uma comparação pixel a pixel, a métrica leva em consideração aspectos de luminância, contraste e estrutura de uma imagem qualquer. Portanto, ao mesmo tempo que fornece uma quantificação mais profunda da qualidade de construção como um todo, ela não vai mensurar diretamente aspectos isolados como geometria e posição.

        In most of these studies, the proposed algorithms are tested and evaluated using metrics such as the mean square error (MSE) of pixel-by-pixel contrast estimation or similar approaches \cite{yin2023subspace,zhang2023iterative,salucci2022learned,wang2023push,liu2022som,bevacqua2021simple,bevacqua2021effective}. This type of approach directly considers the precision of contrast estimation and can indirectly provide perspective on how well the reconstruction of object position and geometry might be, since errors in geometry and positioning can affect this metric. However, a highly accurate geometry with poor contrast estimation can result in considerable error in this type of metric, for example. Therefore, aspects of comparing performance regarding how well object geometry and position are recovered are not adequately addressed by this type of metric. Furthermore, the MSE metric is not suitable for evaluating qualitative algorithms, since these do not estimate the medium's contrast.

        It is also worth highlighting the Structural Similarity (SSIM) metric \cite{wang2004image} to quantitatively show the reconstruction visual quality, which is designed to compute the mismatch between reconstructed contrast results and corresponding ground truths both on the pixel-wise level and perceptual level \cite{wang2023push,liu2025exploring}. Instead of performing a pixel-by-pixel comparison, this metric considers aspects of luminance, contrast, and structure of any given image. Therefore, while it provides a deeper quantification of overall reconstruction quality, it does not directly measure isolated aspects such as geometry and position.

        Specifically for breast imaging, a set of metrics has been proposed in the literature that considers not only the precision in estimating electrical properties, but also the accuracy of the reconstructed tissue geometry \cite{kurrant2021evaluating}. Each geometry (original and reconstructed) is converted into vectors, and the metric is calculated using a formula that performs the dot product and norm of these vectors. To the best of the authors' knowledge, this is the only metric that can isolately consider the geometry aspect of the scatterer, even though it is designed specifically for breast imaging.

@@

        % O objetivo deste trabalho é fornecer mais métricas para auxiliar na avaliação do desempenho de algoritmos. Com uma gama maior de indicadores pode ser possível ter uma visão mais abrangente sobre a capacidade dos algoritmos em diferentes cenários. Por isso, nossa contribuição se baseia na proposição de dois novos indicadores para avaliar o desempenho dos algoritmos na recuperação da geometria e posição dos espalhadores. O primeiro indicador quantifica o erro de forma, enquanto o segundo quantifica o erro de posição. A originalidade do trabalho reside no fato de que estes indicadores podem calcular de maneira mais isolada esses dois aspectos que são relevantes da descrição do desempenho de algoritmos para o problema inverso de espalhamento eletromagnético. Além disso, estes indicadores são projetados para serem aplicáveis a qualquer algoritmo que reconstrua a imagem, independentemente de estimar ou não o contraste.

        The objective of this work is to provide additional metrics to assist in evaluating algorithm performance. With a broader range of indicators, it becomes possible to obtain a more comprehensive understanding of algorithm capabilities across different scenarios. Therefore, our contribution is based on proposing two novel indicators for evaluating algorithm performance in recovering scatterer geometry and position. The first indicator quantifies shape error, while the second quantifies position error.

        The originality of this work lies in the fact that these indicators can measure these two relevant aspects of electromagnetic inverse scattering algorithm performance in a more isolated manner. Furthermore, these indicators are designed to be applicable to any algorithm that reconstructs images, regardless of whether it estimates contrast or not.

@@

        The remainder of this paper is organized as follows. Section \ref{sec:problemstatement} presents the problem statement and the mathematical formulation of the electromagnetic inverse scattering problem. Section \ref{sec:indicators} introduces the proposed indicators for evaluating algorithm performance. Section \ref{sec:results} presents numerical results demonstrating the effectiveness of the proposed indicators. Finally, Section \ref{sec:conclusion} concludes the paper and discusses future work.

@@

		Let $D \in \mathbb{R}^2$ denote the DOI embedded within a homogeneous, isotropic, nonmagnetic ($\mu = \mu_0 = 4\pi \times 10^{-7}$ H/m), and lossless ($\sigma = \sigma_0 = 0$ S/m) background medium with permittivity $\epsilon_b = \epsilon_{rb}\epsilon_0$, where $\epsilon_0 \approx 8.85 \times 10^{-12}$ F/m. We consider a 2D Transverse Magnetic (TMz) polarization where the DOI is illuminated by incident plane waves and assume the time convention $e^{j\omega t}$. All scatterers are located within the DOI, and the scattered field $E^s_z$ at point $\brho \in S$ outside $D$ is evaluated according to the following integral equation \cite{harrington2001time}:
		\begin{equation}
			E^s_z (\brho) = -\frac{k_b^2}{4} \int_D H_0^{(2)}(k_b|\brho-\brhop|)\chi(\brhop)E_z(\brhop) dS^\prime
		\end{equation}

@@

        The experiments were conducted using the open-source library EISPY2D \cite{batista2025eispy2d}, specifically developed to design and evaluate algorithms for the inverse electromagnetic scattering problem. Both indicators will be analyzed, with each approach featuring two case studies (one fundamental and the other exploring a particular aspect) as well as a benchmarking study. The case studies aim to assess performance in specific scenarios, while the benchmarking study facilitates broader generalization of the results. Additionally, an experiment considering a breast imaging scenario was conducted to illustrate the applicability of the proposed indicators in a realistic scenario.

@@

        For all experiments in Section \ref{sec:results:shape} and \ref{sec:results:position}, the following common parameters were used: background wavelength $\lambda_b = 1$ m, noise level of 5\%, relative permittivity of the background medium $\epsilon_{rb} = 1$, and incident wave amplitude of 1 V/m.

@@

                The DNL distribution is shown in Fig. \ref{fig:position:average:dnl}. The values range from below 1 to above 4, with a median close to 2. The high contrast level and geometric variations contributed to the wide distribution of DNL values. However, as shown in Fig. \ref{fig:position:average:zeta_s}, the algorithm performance within each instance remained within a similar range. CSI clearly exhibited the worst performance, followed by SOM. The best performance in each instance varied between OSM and CA. These results are coherent with the findings from Section \ref{sec:results:position:single}.

@@

                % Embora a Fig. \ref{fig:position:average:zeta_s} sugira claramente que há uma diferença no desempenho médio dos algoritmos, foi aplicado novamente o RCBD para fornecer evidências estatísticas para isso. O CSI foi removido desta análise, uma vez que seu desempenho foi significativamente pior que os demais algoritmos. Após validar o pressuposto de normalidade, o valor-p obtido pelo RCBD foi menor que $10^{-30}$. Nas comparações todos-contra-todos feitas através de múltiplos testes T pareados com correção de Bonferroni, o valor-p foi menor que $10^{-6}$ para todos os pares de algoritmos, a hipótese nula de igualdade de desempenho médio só não foi rejeitada para o par OSM-CA. Ou seja, não foram encontradas evidências estatísticas de que OSM e CA possuem desempenhos médios diferentes, sugerindo então que ambos os métodos possuem desempenhos equivalentes. Os valores-p obtidos pelos testes são apresentados na Tabela \ref{tab:position:average:pvalue}.

                While Fig. \ref{fig:position:average:zeta_s} clearly suggests differences in average algorithm performance, RCBD was applied to provide statistical evidence. CSI was excluded from this analysis due to its significantly inferior performance compared to other algorithms.

                After validating the normality assumption, the RCBD yielded a p-value less than $10^{-30}$, indicating significant differences among algorithms. Multiple paired T-tests with Bonferroni correction revealed p-values less than $10^{-6}$ for all algorithm pairs, except for the OSM-CA comparison. The null hypothesis of equal average performance was rejected for all pairs except OSM-CA, indicating that these two methods demonstrate statistically equivalent performance. The complete p-values are presented in Table \ref{tab:position:average:pvalue}.

@@

                Fig. \ref{fig:position:average:confidence_intervals} shows the 95\% confidence intervals for the mean paired differences of the position error indicator $\zeta_P$ for each algorithm pair. The results demonstrate that OSM and CA achieve superior performance compared to BIM and SOM, while BIM outperforms SOM. However, the mean differences do not exceed 1\% in any case, indicating that although statistically significant differences exist, these differences would only be practically relevant in applications requiring high precision in position detection.

@@

            Finally, a case study using a realistic breast phantom demonstrates the application of the proposed indicators in clinical scenarios. This phantom was obtained from the UWCEM Numerical Breast Phantom Repository \cite{burfeindt2012mri} and corresponds to the Very Dense type (ACR class 4).

            Since the repository provides three-dimensional phantoms while our scattering model is two-dimensional, a representative cross-section focusing on the breast interior was extracted, as shown in Fig. \ref{fig:breast:phantom:ground_truth}. The resulting image contains 328 $\times$ 328 pixels with a spatial resolution of 0.5 mm $\times$ 0.5 mm per pixel.

            The phantom includes four tissue types: fatty, transitional, fibroglandular, and skin. The breast is immersed in a homogeneous background medium with relative permittivity $\epsilon_{rb} = 10$ and conductivity $\sigma = 0$ S/m. The operating frequency was set to 1 GHz, with electrical properties of breast tissues determined using the Debye model.

            The experimental configuration follows the previous studies: 80 incident sources, 80 measurement points, and 5\% noise level. Under these conditions, the DNL for this experiment is 0.05. Since the interior of the breast is a complex structure, keeping a low DNL value is an important strategy to facilitate the reconstruction.
